---
title: "NHDD"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

We have two research goals: (1) To investigate the effect of the 2008 recession in the New Haven housing saleprice; (2) To investigate the effects of all the predictor variables in the housing saleprice.  Our initial descriptive statistics focused on the housing prices over the sale dates within each zone. Since we believe that the covariate - "number of bedrooms" contains a lot of the information represented by other covariates such as: number of bathrooms, number of halfbaths, style, and even the square foot of the house, we fixed on this bedrooms variable to look at the trend of the housing saleprice. The scatter plot of the commercial properties (represented by "number of bedrooms" equal to 0) shows no discernible pattern. However, we observe an obvious drop in the housing saleprice during the 2008 recession and even within 1-2 years post-recession. Furthermore, the recovery post-recession is especially slow in the poorer areas, in contrast to the wealthier areas. This conclusion agrees with the data analyzed by The Washington Post which shows that the housing recovery has been strongest in the wealthiest areas, and slowest in the poorest. The average price of a house in a zip code in the top 10 percent of the wealth distribution rose more than 20 percent between 2004 and 2015, versus just 13 percent in the rest. Home values at the high and low ends of the market are now further apart than they've ever seen over this time. In addition, the trends of the housing saleprice at each fixed number of bedrooms stack on top of each other, in the sense that the more the bedrooms that a house has, it appears that the saleprice tends to be higher. Through the use of Linear models regression, Forward Stepwise regression, and Backward Elimination for the overall dataset and for within each zone, we made the following conclusion: (1) Zone, representing the socialeconomic status, has a substantial effect, in explaining the New Haven housing saleprice; (2) A model with the predictor variables: square foot, bedrooms, bathrooms, acres, give the lowest AIC value. 

## Group by neighborhood stuff

```{r}
# library for sample_n
library('dplyr')
library('leaps')
folder <- 'geos'
files <- dir(folder, pattern=".csv")
alldf <- list()
for (i in 1:length(files)) {
  student <- gsub("([^_]*)_.*.csv", "\\1", files[i])
  alldf[[student]] <- read.csv(paste0(folder, "/", files[i]), as.is=TRUE)
}
dd <- read.csv('final_nhh.csv', as.is = TRUE)
d <- sample_n(dd, 5000)
# turn each sale into its own observation
df <- data.frame(matrix(nrow = 1, ncol = 16))
insert.cols <- c('pid', 'yearbuilt', 'totval', 'bedrooms', 'bathrooms', 'halfbaths', 'landval', 'zone', 'acres', 'exval', 'pctgood', 'sqft', 'garagesqft', "saleprice", "saledate", "saleowner")
colnames(df) <- insert.cols
for (i in seq(1:nrow(d)) ) {
  # for each sale in row
  for (sale in 1:5) {
    # create new row with the right data
    access.cols <- c('pid', 'yearbuilt', 'totval', 'bedrooms', 'bathrooms', 'halfbaths', 'landval', 'zone', 'acres', 'exval', 'pctgood', 'sqft', 'garagesqft', paste(c("saleprice", "saledate", "saleowner"), sale, sep = ""))
    
    r_ <- d[i, access.cols]
    names(r_)
    
    colnames(r_) <- insert.cols
    df <- rbind(df, r_)
    
  }
}
# turn the dates into separate cols via lubridate
library('lubridate')
lbd_helper1 <- mdy(df$saledate)
df$salemonth <- as.factor(month(lbd_helper1))
df$saleyear <- year(lbd_helper1)
```

Data cleaning:
```{r}
# drop the na sales ones
df.na <- df[-which(is.na(df$saledate)),]
df.na$saleprice <- as.numeric(df.na$saleprice)
# drop the ones in which the sale price was 0
df.sale <- df.na[-which(df$saleprice == 0), ]
```

Adding the lat and longitude data
```{r}
# add latitude and longitude data
adf <- data.frame(matrix(ncol=3))
colnames(adf) <- c('pid', 'latitude', 'longitude')
for (d in alldf) {
  newdf <- data.frame(d$pid)
  colnames(newdf) <- c('pid')
  newdf$latitude <- ifelse('latitude' %in% colnames(d), d$latitude, d$lat)
  newdf$longitude <- ifelse('longitude' %in% colnames(d), d$longitude, d$long)
  adf <- rbind(adf, newdf)
}
# some people messed up the lat's and long's 
# get all the ones where the latitude is 
df.loc <- merge(adf, df.sale, by = 'pid')
```

separate into res vs com:
```{r}
# separate into commercial vs residential units
df.com <- df.loc[which(df.loc$bedrooms == 0), ]
df.res <- df.loc[which(df.loc$bedrooms != 0), ]
# get rid of all property transfers
df.res.s <- df.res[-which(df.res$saleprice == 0), ]
```

## Analyzing the change in sale prices by neighborhood
```{r}
# get average price for zone by each year
# dataframe to store stuff
avg.nghbd.price <- data.frame(
  matrix(
    ncol = length(unique(df.res.s$saleyear)),
    nrow = length(unique(df.res.s$zone))
    )
  )
colnames(avg.nghbd.price) <- as.character(unique(df.res.s$saleyear))
rownames(avg.nghbd.price) <- unique(df.res.s$zone)
for (zone in unique(df.res.s$zone) ) {
  for (year in unique(df.res.s$saleyear)) {
    m <- mean(df.res.s$saleprice[which(df.res.s$saleyear == year & df.res.s$zone == zone)], na.rm = TRUE)
    avg.nghbd.price[zone, as.character(year)] <- m
  }
  
}
```
Okay, now for some graphs
```{r}
plot(log(avg.nghbd.price['RS2',]))
```

## Counfounding parameters
- Zoning law changes
- Immigration